\section{Objectives}
\label{sec:objetivos}

\subsection*{General Objective}

To develop an interactive, agent-guided simulation platform that enables users to explore and understand orbital mechanics through embodied interaction, combining physical manipulation, natural language dialogue, and real-time visualizations.

\subsection*{Specific Objectives}

\begin{enumerate}
\item Design and implement a simulation environment capable of rendering orbital trajectories in real time, grounded in physically accurate models.
\item Integrate a generative agent capable of interpreting natural language input, translating it into simulation parameters, and guiding the user through explanations and interactions.
\item Develop a physical interface, centered around a globe equipped with sensors, that captures real-world rotational input and transmits it into the simulation system.
\item Enable multimodal interaction by combining physical movement, voice commands, and visual feedback to create a seamless and intuitive user experience.
\item Ensure that all components of the system, simulation, agent, and physical bridge, function coherently and communicate reliably in real time.
\item Create a system architecture that is modular and extensible, allowing for future expansion to other celestial bodies, educational modules, or mission types.
\item Evaluate the platform’s potential as an educational tool for facilitating conceptual understanding of orbital mechanics through exploratory learning.
\end{enumerate}